BIP-Predictor

### 2025-05-23 Sweave/BIP_predictor.Rnw
up

#------------------------------
# Section 2.1: data
recompute_results<-F

@@

# Settings for M-SSA
ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
f_excess<-c(5,rep(4,length(select_vec_multi)-1))
# In-smaple span for VAR
date_to_fit<-"2008"
# We can select type of VAR model
VAR_type="BVAR"
VAR_type="VAR"
if (VAR_type=="BVAR")
{
# Simon's settings for BVAR(4)  
  lambda_BVAR <- 0.001
  p<-3
  q<-0
}
if (VAR_type=="VAR")
{
# Marc's settings for VAR  
  p<-1
  q<-0
}

@@

    set.seed(12)
    V_obj<-VARMA(data_fit,p=p,q=q)
# Apply regularization: see vignette MTS package
    threshold<-1.5
    V_obj<-refVARMA(V_obj, thres = threshold)
    save(V_obj_wc,file=paste(path.result,"/VAR_wc",sep=""))  
  }
  if (VAR_type=="BVAR")
  {
    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
    #   low values <=> low precision
    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
    V_obj_wc <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
    save(V_obj_wc,file=paste(path.result,"/BVAR_wc",sep=""))
  }

@@

# The following double loop computes all combinations of forward-shifts (of BIP) and forecast horizons (of M-SSA)
  for (shift in shift_vec)#shift<-1
  {
# Progress bar: see R-console
    setTxtProgressBar(pb, shift)
    for (j in h_vec)#j<-1
    {
# Horizon j corresponds to k=j+1-th entry of array    
      k<-j+1
# A. M-SSA component predictor
# Specify data matrix for WLS regression
      if (length(sel_vec_mssa_comp)>1)
      {
        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),t(mssa_array[sel_vec_mssa_comp,,k]))
      } else
      {
        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),(mssa_array[sel_vec_mssa_comp,,k]))
      }
      rownames(dat)<-rownames(x_mat)
      colnames(dat)<-c(colnames(x_mat)[1],sel_vec_mssa_comp)
      nrow(na.exclude(dat))

@@

if (length(sel_vec_mssa_comp)>1)
{
  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),t(final_mssa_array[sel_vec_mssa_comp,,h+1]))
} else
{
  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),(final_mssa_array[sel_vec_mssa_comp,,h+1]))
}
lm_obj<-lm(dat[,1]~dat[,2:ncol(dat)])

optimal_weights<-lm_obj$coef
# Compute predictor for each forward-shift  
if (length(sel_vec_mssa_comp)>1)
{
  final_mssa_predictor<-optimal_weights[1]+dat[,2:ncol(dat)]%*%optimal_weights[2:length(optimal_weights)]
} else
{
  final_mssa_predictor<-optimal_weights[1]+dat[,2:ncol(dat)]*optimal_weights[2:length(optimal_weights)]
}

>>>>

#------------------------------
# Section 2.1: data
recompute_results<-T

@@

# Settings for M-SSA
ht_mssa_vec<-c(6.380160,  6.738270,   7.232453,   7.225927,   7.033768)
f_excess<-c(5,rep(4,length(select_vec_multi)-1))
# In-smaple span for VAR
date_to_fit<-"2008"
# We can select type of VAR model
VAR_type="BVAR"
#VAR_type="VAR"
if (VAR_type=="BVAR")
{
# Simon's settings for BVAR(4)  
  lambda_BVAR <- 1.001
  p<-3
  q<-0
}
if (VAR_type=="VAR")
{
# Marc's settings for VAR  
  p<-1
  q<-0
}

@@

    set.seed(12)
    V_obj<-VARMA(data_fit,p=p,q=q)
# Apply regularization: see vignette MTS package
    threshold<-1.5
    V_obj_wc<-refVARMA(V_obj, thres = threshold)
    save(V_obj_wc,file=paste(path.result,"/VAR_wc",sep=""))  
  }
  if (VAR_type=="BVAR")
  {
    V0_BVAR <- diag(rep(1,n))  # prior for Sigma_a matrix - K x K
    # precision matrix of size (kp+1) x (kp+1)  if there's a constant
    #   low values <=> low precision
    C_BVAR <- lambda_BVAR*diag(rep(1,n*p+1))
    V_obj_wc <- BVAR(data_fit,p = p,C = C_BVAR,V0 = V0_BVAR)
    save(V_obj_wc,file=paste(path.result,"/BVAR_wc",sep=""))
  }

@@

# The following double loop computes all combinations of forward-shifts (of BIP) and forecast horizons (of M-SSA)
  for (shift in shift_vec)#shift<-4
  {
# Progress bar: see R-console
    setTxtProgressBar(pb, shift)
    for (j in h_vec)#j<-4
    {
# Horizon j corresponds to k=j+1-th entry of array    
      k<-j+1
# A. M-SSA component predictor
# Specify data matrix for WLS regression
      if (length(sel_vec_mssa_comp)>1)
      {
        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),t(mssa_array[sel_vec_mssa_comp,,k]))
      } else
      {
        dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),(mssa_array[sel_vec_mssa_comp,,k]))
      }
      rownames(dat)<-rownames(x_mat)
      colnames(dat)<-c(colnames(x_mat)[1],sel_vec_mssa_comp)
      nrow(na.exclude(dat))

@@

if (length(sel_vec_mssa_comp)>1)
{
  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),t(final_mssa_array[sel_vec_mssa_comp,,h+1]))
} else
{
  dat<-cbind(c(x_mat_wc[(shift+lag_vec[1]+1):nrow(x_mat_wc),1],rep(NA,shift+lag_vec[1])),(final_mssa_array[sel_vec_mssa_comp,,h+1]))
}
if (reg_type=="OLS")
{
  lm_obj<-lm(dat[,1]~dat[,2:ncol(dat)])
}  
if (reg_type=="Ridge")
{
  lm_obj<-lm.ridge(dat[,1]~matrix(cbind(rep(1,ncol(dat)-1),dat[,2:ncol(dat)]),ncol=ncol(dat))-1, lambda = lambda_reg)
}
if (reg_type=="Lasso")
{
  lm_obj<-glmnet(x=dat[,2:ncol(dat)],y=dat[,1],alpha=0,lambda = lambda_reg)
}
optimal_weights<-lm_obj$coef
# Compute predictor for each forward-shift  
if (length(sel_vec_mssa_comp)>1)
{
  final_mssa_predictor<-optimal_weights[1]+dat[,2:ncol(dat)]%*%optimal_weights[2:length(optimal_weights)]
} else
{
  final_mssa_predictor<-optimal_weights[1]+dat[,2:ncol(dat)]*optimal_weights[2:length(optimal_weights)]
}
